\section{Some Basics Definitions of Integers}


\input{numberTheory/examples/evenOdd.tex}
\input{numberTheory/examples/primeComposite.tex}

\input{numberTheory/examples/constructiveProofs.tex}

Direct proofs:

\input{numberTheory/props/sumOfEvenIsEven.tex}

\input{numberTheory/exercises/sumOfOddIsEven.tex}
\input{numberTheory/exercises/sumOfEvenAndOddIsOdd.tex}
\input{numberTheory/exercises/productOfOddIsOdd.tex}
\input{numberTheory/exercises/productOfEvenAndOddIsEven.tex}
\input{numberTheory/exercises/showExpressionIsInteger.tex}

\input{numberTheory/examples/someRationals.tex}
\input{numberTheory/props/integersAreRational.tex}

\input{numberTheory/props/doubleOfRationalIsRational.tex}
\input{numberTheory/props/productOfRaionalsIsRational.tex}


\input{numberTheory/examples/everyIntegerDivides0.tex}

\input{numberTheory/props/divisorsOf1.tex}
\input{numberTheory/props/everyPositiveNonUnitIntegerDivisibleByAPrime.tex}

The proof of this next lemma requires a technique which we have not covered, yet.
This techniques is called strong induction.
Strong induction is a strengthening of regular induction.
Regular induction is a proof technique to show a claim holds over a well-ordered sets (which do not contain a limit -- this is outside the scope of this class).
You show that a claim holds at a base cases.
After the base case you show the successor case; that if the claim holds at a specific place, then it necessarily holds at the next place.
After you show these two things,
Strong induction differs from regular, weak, induction in the successor case.
In strong induction, we suppose that the claim holds everywhere at and bellow a point and then show that it holds at the next point.

The above exposition is not necessary.
Though, as induction is one of the more valuable skills learned in this course, I want to get a little jump start the the subject.
\input{numberTheory/lems/positiveNotUnitIntegerPrimeOrProductOfPrimes.tex}

The proof of the next theorem requires the use of the fact that $\NN$ is well-ordered by $<$.
This means two things.
First, $\NN$ is totally ordered by $<$, that is $\forall a,b\in\NN$ $a<b\vee a=b\vee b<a$.
Second, every nonempty subset of $\NN$ has a $<$ least element.
\input{numberTheory/thms/fundementalTheoremOfArithmetic.tex}

% \input{numberTheory/thms/quotientRemainderTheorem.tex}
