\section{Some Proofs Worth Knowing}

The following section contains some classic Theorems the proofs of which every young mathematician is expected to be familiar with.

\input{numberTheory/thms/rootTwoIrrational.tex}

Legend holds that the previous Theorem cost the life of a mathematician thousands of years ago.
The legend, as I was taught, is as follows:\\
The Pythagoreans, those that follow the philosophy Pythagoras, believed that the universe was built of whole numbers.
What we call integer.
Further, they believed that the universe, and every measurement within it, is rational.

Now, the name Pythagoras aught to be familiar, as that is the name which we attribute the following well know geometric theorem:
\input{geometry/thms/PythagoreanTheorem.tex}
The Pythagorean Theorem, Theorem \ref{thm:PythagoreanTheorem}, can be proved using a straight-edge and compass.
Further, the object which the Pythagorean Theorem applies can be construct, in a real way.
Thus, the measurement given by the Pythagorean Theorem truly exist.
So, if one draws a unit square, whatever that unit may be, then the diagonal of that square must necessarily square to be 2 units.
That is, the square root necessarily exists within our world.

The consequences of Theorem \ref{thm:rootTwoIrrational}, when presented to them, were catastrophic to the world view of the Pythagoreans.
Necessarily, their world view was, provably, incorrect.
So, when Hippasus of Metapontum presented this theorem to his fellow Pythagoreans, his fellow Pythagoreans drowned him.

\input{numberTheory/thms/infinitelyManyPrimes.tex}
