\section{Suppose Otherwise}

\input{logic/defns/tautology.tex}

The central idea of proving a statement by contradiction is that mathematics must be non-contradictory.
That is, for any statement $\phi$ either $\phi$ is true or $\phi$ is false but not both!
More formally, the statement form $\neg(p\wedge\neg p)$ is a tautology.

It should be noted that proofs by contrition, while valid, are often undesirable.
This is because proofs by contradiction often do not reveal the same insight offered by constructive or direct proofs.
Further, proofs by contradiction can often, if done without care, feel inelegant, much like a sledgehammer.

Suppose we are attempting to prove a statement $\phi$ via contradiction.
The common structure of a proof by contradiction is as follows.
\begin{proof}
  Suppose, towards a contradiction, $\neg\phi$.\\
  Unpack what $\neg\phi$ means.\\
  Perform some work and obtain a formula $\neg\psi$ where we know $\psi$ to be true before we supposed $\neg\phi$.\\
  This gives $\psi$ and $\neg\psi$, a contradiction.\\
  Whence, $\phi$.
\end{proof}
Now, the above is a simple sketch of a near ideal argument.
In practice, this argument can take many forms.
In this section, we will show some examples of this technique to help the reader become more familiar with the process.

\input{numberTheory/props/integersAreUnbounded.tex}
\input{numberTheory/props/noIntegerBothEvenAndOdd.tex}
\input{numberTheory/props/sumOfRationalAndIrrationalIsIrrational.tex}

% We need the following for the next discussion.
\input{logic/defns/contrapositive.tex}
\input{logic/defns/universalConditionalStatement.tex}

Proof by contrapositive.\\
Suppose we are tasked with proving a statement of the form $\forall x( P(x)\rightarrow Q(x))$.
Suppose further that the predicate $P(x)$ gives little in the way of describing $Q(x)$ (that is, showing $\forall x( P(x)\rightarrow Q(x))$ is hard).
Then, maybe before we attempt a proof via contradiction, we can attack the problem directly from a different direction.
That is, recall that $\forall x( P(x)\rightarrow Q(x))$ is logically equivalent to its contrapositive $\forall x( \neg Q(x)\rightarrow \neg P(x))$.
So, a proof of $\forall x( P(x)\rightarrow Q(x))$ by contradiction is a direct proof of $\forall x( \neg Q(x)\rightarrow \neg P(x))$.
The following lemma is an example of this technique.
\input{numberTheory/lems/IfSquareOfIntegerIsEvenThenEven.tex}
