\guard

\input{setTheory/props/integersAreCountable.tex}
\input{setTheory/thms/schroderBernstein.tex}
\input{setTheory/props/naturalsCrossNaturalsisCountable.tex}

\begin{prop}
\label{prop:rationalsAreCountable}
  The set of rationals, $\QQ$, is countable.
\end{prop}
\begin{proof}
  We will show this in steps.
  Assume that we have constructed a bijection $f:\ZZ^+\to \QQ^+$.
  Note this map, after multiplying by $-1$, sees that negative rationals are have the same cardinality as $\ZZ$ as well.
  Now, with this bijection, we can form a bijection $g:\ZZ\to\QQ$ defined as as \[ g(z) = \begin{cases}
      0 & z= 0 \\
      -f(-z) & z<0 \\
      f(z) & z>0
   \end{cases} \]
   for each $z\in\ZZ$.
   With this bijection in hand, we use the fact that $\ZZ$ is countable, Proposition \ref{prop:integersAreCountable}, together with Proposition \ref{prop:cardinalityIsTransitive} to obtain that $\QQ$ is countable.

   So, it remains to show that there is a bijection $f:\ZZ^+\to\QQ^+$.
   (There is a classic zig-zag argument which informally describes such a bijection. This zig-zagging can be made formal, and a closed form of that description can be written. I will present that in lecture, but for the notes I will show a different way.)
   As $\ZZ^+\subset \QQ^+$, there exists, trivially, an injection $\ZZ^+\to\QQ^+$, say $z\mapsto \frac{z}{1}$.
   Though, as $\vert \ZZ^+\vert=\vert\NN\vert$, we can produce an injection from $\NN$ to $\QQ$, say $n\mapsto \frac{n+1}{1}$.

   Define an injection from $\QQ^+$ to $\NN$ as follows
   First, we define $h:\QQ^+\to\NN\times\NN$.
   For each $r\in\QQ$, write $r=\frac{a}{b}$ where $\frac{a}{b}$ is completely reduced.
   Define $h(r)=(a,b)$.
   Certainly, this map is an injection.
   We use the fact that $\vert\NN\times\NN\vert=\vert\NN\vert$ to obtain a bijection $i:\NN\times\NN\to\NN$.
   Then $i\circ h$ is the desired injection.

   Finally, applying Schroder-Bernstein Theorem, we have that there exists a bijection $f':\NN\to\QQ^+$.
   Set $f:\ZZ^+\to\QQ^+$ to be $f'(n+1)$.
\end{proof}
