\guard

\section{Predicates and Quantifiers}
\label{sec:predicatesAndQuantifiers}

Note the sentence
  \begin{center}
    ``They are a college student.''
  \end{center}
  is not a statement as it may be true or false depending on the value of ``They''.
In this sentence ``They'' is a free variable.
Similarly, ``$x+y>0$'' is not a statement, it has more familiar variables $x$ and $y$.

The predicate refers to the part of the sentence with some or all of its nouns removed.
That is, in the sentence
  \begin{center}
    ``James is a student at the University of North Texas.'',
  \end{center}
  ``James'' is the subject and
  \begin{center}
    ``is a student at the University of North Texas.''
  \end{center}
  is the predicate.

In logic, predicates are formed in much the same way.
\input{logic/examples/predicates.tex}
In Example \ref{exmp:predicates}, we referred to the example's predicates with their full names, $P(x)$ and $Q(x,y)$.
When there is no room for confusion and the variables or other ornaments are of little importance to the statement being made, we may drop those decorations from the symbol.
That is, the concluding sentence of Example \ref{exmp:predicates} could be written ``Whence, $P$ and $Q$ are predicates.''.


\input{logic/examples/truthSetOverReals.tex}

\input{logic/examples/truthSetOverIntegers.tex}


\input{logic/examples/univeralStatementOverFiniteSet.tex}

\index{method of exhaustion}
Note that the technique of checking each element of the domain of an universal statement, as in the example above, is not the way to go.
This method is called the \emph{method of exhaustion}, and is aptly named.

\input{logic/examples/univeralStatementOverRealsFalse.tex}
The technique of producing a counter example is the standard way to show that  an universal statement is false.


\input{logic/examples/existentialStatementOverFiniteSetFalse.tex}
\input{logic/examples/existentialStatementOverReals.tex}

\input{logic/examples/translateQuantifiedStatementsFromEnglish.tex}

\input{logic/examples/translateQuantifiedConditionalStatements.tex}

Note that the statements
\begin{center}
	``For all real numbers $x$ if $x$ is an integer, then $x$ is a rational.''
\end{center}
and
\begin{center}
	``For all integers $x$, $x$ is a rational.''
\end{center}
are equivalent.

In fact, \[\forall x\in D(P(x)\rightarrow Q(x) )\equiv \forall x\in \set{y\in D\mid P(x)}( Q(x) )\]


Let $P(x)$ and $Q(x)$ be predicates with $D$ the common domain of $x$.
We write $P(x)\implies Q(x)$ to mean $\forall x\in D( P(x)\rightarrow Q(x))$ when there is no risk for confusion of the domain $D$.
Similarly, we write $P(x)\iff Q(x)$ to mean $\forall x\in D( P(x)\leftrightarrow Q(x))$

\input{logic/exercises/negationOfExistentialStatement.tex}
\input{logic/examples/negateQuantifiedStatement.tex}

\input{logic/exercises/negationOfUniversalConditionalStatement.tex}
\input{logic/examples/negationOfUniversalConditionalStatement.tex}

Recall that we say that a conditional statement, $p\rightarrow q$ is vacuously true provided $p$ is false.
In the same way, we say that a statement of the form $\forall x\in D(P(x)\rightarrow Q(x))$ is vacuously true if $\forall x\in D$ $P(x)$ is true.
Further, we define the converse, inverse, and contrapositive fir quantified conditional statements similar to how we defined them for conditional statements $p\rightarrow q$.
That is, for a statement of the form $P(x)\implies Q(x)$:
\begin{itemize}
  \item its converse is $Q(x)\implies P(x)$,
  \item its inverse is $\neg P(x) \implies \neg Q(x)$, and
  \item its contrapositive is $\neg Q(x) \implies \neg P(x)$.
\end{itemize}

\begin{exercise}
  Show that $P(x)\implies Q(x)$ is logically equivalent to $\neg Q(x)\implies \neg P(x)$.
\end{exercise}

\begin{exercise}
  Show that $P(x)\implies Q(x)$ is not logically equivalent to $Q(x)\implies P(x)$ and $\neg P(x)\implies \neg Q(x)$.
\end{exercise}

Similar to Definition \ref{defn:sufficentAndNecessary}
\begin{itemize}
  \item ``For all $x$, $R(x)$ is a sufficient condition for $S(x)$'' means $R(x)\implies S(x)$.
  \item ``For all $x$, $R(x)$ is a necessary condition for $S(x)$'' means $\neg R(x)\implies \neg Q(x)$.
  \item ``For all $x$, $R(x)$ only if $S(x)$'' means $\neg S(x)\implies \neg R(x)$.
\end{itemize}
